Adaptive Approximation for Multivariate Linear Problems with Inputs Lying in a Cone

### 2019-02-13 DingHickKritzMakRICAMChap.tex
Fixed Sect. 3 optimality proof I think

\usepackage[utf8]{inputenc}%(only for the pdftex engine)
%\RequirePackage[no-math]{fontspec}[2017/03/31]%(only for the luatex or the xetex engine)
\usepackage[small]{dgruyter}
\usepackage{microtype}
\usepackage{amsmath,amssymb}
\usepackage{subcaption}
\input{macros}

@@

\theoremstyle{dgthm}
\newtheorem{theorem}{Theorem}
\theoremstyle{dgthm}
\newtheorem{lemma}{Lemma}
\theoremstyle{dgthm}
\newtheorem{corollary}{Corollary}
\theoremstyle{dgthm}
\newtheorem{proposition}{Proposition}
\theoremstyle{dgdef}
\newtheorem{remark}{Remark}

\DeclareMathOperator{\SOL}{SOL}
\DeclareMathOperator{\APP}{APP}
\DeclareMathOperator{\ALG}{ALG}
\DeclareMathOperator{\ERR}{ERR}
\newcommand{\dataN}{\bigl(\hf(\bsk_i)\bigr)_{i=1}^n}
\newcommand{\ERRN}{\ERR\bigl(\dataN,n\bigr)}
\DeclareMathOperator{\COST}{COST}
\DeclareMathOperator{\COMP}{COMP}
\newcommand{\hf}{\widehat{f}}
\newcommand{\hg}{\widehat{g}}
\newcommand{\tu}{\Tilde{u}}
\newcommand{\tv}{\Tilde{v}}
\newcommand{\tcalK}{\widetilde{\calK}}
\newcommand{\tlambda}{\Tilde{\lambda}}
\newcommand{\tbbK}{\widetilde{\bbK}}
\def\abs#1{\ensuremath{\left \lvert #1 \right \rvert}}
\newcommand{\normabs}[1]{\ensuremath{\lvert #1 \rvert}}
\newcommand{\bigabs}[1]{\ensuremath{\bigl \lvert #1 \bigr \rvert}}
\newcommand{\Bigabs}[1]{\ensuremath{\Bigl \lvert #1 \Bigr \rvert}}
\newcommand{\biggabs}[1]{\ensuremath{\biggl \lvert #1 \biggr \rvert}}
\newcommand{\Biggabs}[1]{\ensuremath{\Biggl \lvert #1 \Biggr \rvert}}
\newcommand{\norm}[2][{}]{\ensuremath{\left \lVert #2 \right \rVert}_{#1}}
\newcommand{\normnorm}[2][{}]{\ensuremath{\lVert #2 \rVert}_{#1}}
\newcommand{\bignorm}[2][{}]{\ensuremath{\bigl \lVert #2 \bigr \rVert}_{#1}}
\newcommand{\Bignorm}[2][{}]{\ensuremath{\Bigl \lVert #2 \Bigr \rVert}_{#1}}
\newcommand{\biggnorm}[2][{}]{\ensuremath{\biggl \lVert #2 \biggr \rVert}_{#1}}
\newcommand{\ip}[3][{}]{\ensuremath{\left \langle #2, #3 \right \rangle_{#1}}}
\newcommand{\FredNote}[1]{{\color{blue}Fred: #1}}
\newcommand{\YuhanNote}[1]{{\color{magenta}Yuhan: #1}}
\newcommand{\PeterNote}[1]{{\color{orange}Peter: #1}}
\newcommand{\SimonNote}[1]{{\color{purple}Simon: #1}}
\newcommand{\tLambda}{\widetilde{\Lambda}}
%\newcommand{\HickernellFJ}{Hickernell}

@@

  \author[1]{Yuhan Ding}
  \author*[2]{Fred J. Hickernell}
  \author[3]{Peter Kritzer} 
  \author[4]{Simon Mak}
  \runningauthor{Y. Ding, F. J. Hickernell, P. Kritzer, and S. Mak}
  \affil[1]{Department of Mathematics, 
  Misericorida University, 301 Lake Street, Dallas, PA 18704 USA}
  \affil[2]{Department of Applied Mathematics, Illinois Institute of Technology, RE 220, 10 W.\ 32${}\text{nd}$ Street, Chicago, IL 60616 USA}
  \affil[3]{Johann Radon Institute for Computational and Applied Mathematics (RICAM), 
  Austrian Academy of Sciences, Altenbergerstr. 69, 4040 Linz, Austria}
  \affil[4]{...}
  \title{Adaptive Approximation for Multivariate Linear Problems with Inputs Lying in a Cone}
  \runningtitle{Adaptive Approximation}
  \subtitle{...}
  \abstract{...}
  %\keywords{...}
  %\classification[PACS]{...}
  %\communicated{...}
  %\dedication{...}
  %\received{...}
  %\accepted{...}
  %\journalname{...}
  %\journalyear{...}
  %\journalvolume{..}
  %\journalissue{..}
  \startpage{1}
  \aop
  %\DOI{...}

@@

In many situations, adaptive algorithms can be rigorously shown to perform \emph{essentially no better} than non-adaptive algorithms.  Yet, in practice adaptive algorithms are appreciated because they relieve the user from having to stipulate the computational effort required to achieve the desired accuracy.  Adaptive algorithms infer the computational effort required based on the function data sampled.  The key to resolving this seeming contradiction is to construct a theory based on assumptions that favor adaptive algorithms.  That is what we do here.

Adaptive algorithms may perform better if the set of input functions is not convex, and that is the situation studied here. We propose adaptive algorithms for general multivariate linear problems where the input function lies in non-convex cones.  Our algorithms use series coefficients of the input function to construct an approximate solution that satisfies an absolute error tolerance.  We show our algorithms to be essentially optimal.  We derive conditions under which the problem is tractable.  In the remainder of this section we define the problem and essential notation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{General Linear Problem}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Our general linear problem is defined by a solution operator that mapping an input function to an output, $\SOL:\calF \to \calG$.  The Banach spaces of inputs and outputs are defined by series expansions:
\begin{gather}
    \calF = \left \{f = \sum_{\bsk \in \mathbb{K}} \hf(\bsk) u_{\bsk} : \norm[\calF]{f} : = \norm[\rho]{\left( \frac{\hf(\bsk)}{\lambda_{\bsk}} \right)_{\bsk \in \mathbb{K}}} < \infty \right\}, \quad 1 \le \rho \le \infty, \\
    \calG = \left \{g = \sum_{\bsk \in \mathbb{K}} \hg(\bsk) v_{\bsk} : \norm[\calG]{g} : = \norm[\tau]{\bigl(  \hg(\bsk)  \bigr)_{\bsk \in \mathbb{K}}} < \infty \right\}, \quad 1 \le \tau \le \rho.
\end{gather}
Here, $\{u_{\bsk}\}_{\bsk \in \mathbb{K}}$ is a basis for the input Banach space $\calF$, and $\{v_{\bsk}\}_{\bsk \in \mathbb{K}}$ is a basis for the output Banach space $\calG$. These bases are defined to match the solution operator:
\begin{equation} \label{DHKM:basis_relate}
    \SOL(u_{\bsk}) = v_{\bsk} \qquad \forall \bsk \in \mathbb{K},
\end{equation}
where $\mathbb{K}$ is a general index set. The $\lambda_{\bsk}$ represent the importance of the series coefficients of the input function.

Although this problem formulation is quite general in some aspects, condition \eqref{DHKM:basis_relate} may be somewhat restrictive.  In theory, the choice of basis can be made via the singular value decomposition, but in practice, if $\calF$ and $\calG$ have common norms, bases satisfying \eqref{DHKM:basis_relate} may be difficult to identify explicitly.

To facilitate our derivations below, we establish the following lemma:

\begin{lemma} \label{DHKM:Key_Lem}
Let $\calK$ be some proper or improper subset of $\bbK$. Moreover, let $\rho'$ be defined by the relation
\begin{equation*}
    \frac 1\rho + \frac 1 {\rho'} = \frac 1 \tau, \qquad \text{i.e., } \rho' = \frac{\rho \tau}{\rho - \tau},
\end{equation*}
so $\tau \le \rho' < \infty$.  Let $\Lambda =  \bignorm[\rho']{\bigl(  \lambda_{\bsk}  \bigr)_{\bsk \in \calK}}$.  Then the following are true for $f = \sum_{\bsk \in \calK} \hf(\bsk) u_{\bsk}$:
\begin{gather}
\label{DHKM:SOL_ineq}
    \norm[\calG]{\SOL(f)} \le \norm[\calF]{f} \, \Lambda,
    \\
    \label{DHKM:SOL_tight_ineq}
    \bigabs{\hf(\bsk)} = \frac{R \lambda_{\bsk}^{\rho'/\rho + 1}}{\Lambda^{\rho'/\rho}} \quad \forall \bsk \in \calK \ \ 
    \implies  \ \ \norm[\calF]{f} = R \ \& \ \norm[\calG]{\SOL(f)} = R \Lambda.
    \end{gather}
Equality \eqref{DHKM:SOL_tight_ineq} illustrates how inequality \eqref{DHKM:SOL_ineq} may be made tight.
\end{lemma}
\begin{proof}
The proof of inequality \eqref{DHKM:SOL_ineq} proceeds by applying H\"older's inequality:
\begin{align}
    \label{DHKM:SOL_A}
    \norm[\calG]{\SOL(f)}  
    & = \biggnorm[\calG]{\sum_{\bsk \in \calK} \hf(\bsk) v_{\bsk}} = \norm[\tau]{\bigl(  \hf(\bsk)  \bigr)_{\bsk \in \calK}} \\
\nonumber
    & = \left [\sum_{\bsk \in \calK}  \left\lvert\frac{\hf(\bsk)}{\lambda_{\bsk}} \right\rvert^{\tau} \lambda_{\bsk}^{\tau} \right]^{1/\tau} \\
    \nonumber
    & \le \biggnorm[\rho]{\biggl(  \frac{\hf(\bsk)}{\lambda_{\bsk}}  \biggr)_{\bsk \in \calK}} \, \bignorm[\rho']{\bigl(  \lambda_{\bsk}  \bigr)_{\bsk \in \calK}} \qquad \text{since }\frac 1\rho + \frac 1 {\rho'} = \frac 1 \tau \\
    \nonumber
    & = \norm[\calF]{f} \, \Lambda.
\end{align}
Substituting the formula for $\bigabs{\hf(\bsk)}$ in \eqref{DHKM:SOL_tight_ineq} into equation \eqref{DHKM:SOL_A} and applying the relationship between $\rho$, $\rho'$, and $\tau$ yields
\begin{equation*}
       \norm[\calG]{\SOL(f)}  
    =  \frac{R \bignorm[\tau]{\bigl(  \lambda_{\bsk}^{\rho'/\rho + 1}  \bigr)_{\bsk \in \calK}}} {\Lambda^{\rho'/\rho}} 
    = \frac{R \bignorm[\rho']{\bigl(  \lambda_{\bsk}  \bigr)_{\bsk \in \calK}}^{\rho'/\rho + 1}}
    {\Lambda^{\rho'/\rho}} = R \Lambda.
\end{equation*}
Moreover,
\begin{equation*}
    \norm[\calF]{f}  
    = \norm[\rho]{\left( \frac{\hf(\bsk)}{\lambda_{\bsk}} \right)_{\bsk \in \calK}}
    = \frac{R \bignorm[\rho]{\bigl(  \lambda_{\bsk}^{\rho'/\rho}  \bigr)_{\bsk \in \calK}}}{\Lambda^{\rho'/\rho}} 
    = \frac{R \bignorm[\rho']{\bigl(  \lambda_{\bsk}  \bigr)_{\bsk \in \calK}}^{\rho'/\rho}}
    {\Lambda^{\rho'/\rho}} = R.
\end{equation*}
This completes the proof.
\end{proof}

\hspace{\parindent} Based on this lemma and taking $\calK = \bbK$, the norm of the solution operator can be expressed in terms of the $\lambda_\bsk$ as follows:
\begin{equation} \label{DHKM:SOLNorm}
    \norm[\calF \to \calG]{\SOL}  = \sup_{\norm[\calF]{f} \le 1} \norm[\calG]{\SOL(f)} = \bignorm[\rho']{\bigl(  \lambda_{\bsk}  \bigr)_{\bsk \in \mathbb{K}}}.
\end{equation}
The $\lambda_{\bsk}$ define the smoothness of the input space and have a known order:
\begin{equation} \label{DHKM:lambda_order}
    \lambda_{\bsk_1} \ge \lambda_{\bsk_2} \ge \cdots >0.
\end{equation}
We assume that all $\lambda_{\bsk_i}$ are positive to avoid the trivial case where the problem can be solved by knowing a finite number of series coefficients.
An optimal approximation  based on $n$ series coefficients of the input function can be defined in terms of the series coefficients of the input function corresponding to the largest $\lambda_{\bsk}$ as follows:
\begin{equation} \label{DHKM:APP_def}
    \APP : \calF \times \N_0 \to \calG, \qquad  \APP(f,n) := \sum_{i=1}^n \hf(\bsk_i) v_{\bsk_i}, \quad \APP(f,0) = 0.
\end{equation}

@@

 Let $\APP'(0,n)$ depend on the series coefficients indexed by $\calJ  = \{\bsk'_1, \ldots, \bsk'_n\}$.  Use Lemma \ref{DHKM:Key_Lem} to choose $f$ to satisfy the following conditions:
\begin{gather*}
    \hf(\bsk'_1) = \cdots = \hf(\bsk'_n) = 0, \qquad \norm[\calF]{f} = R, \\ \Bignorm[\tau]{\bigl(\hf(\bsk)\bigr)_{\bsk \notin \calJ}}
    = \norm[\rho]{\left( \frac{\hf(\bsk)}{\lambda_{\bsk}} \right)_{\bsk \notin \calJ}} \,
    \norm[\rho']{\left( \lambda_{\bsk} \right)_{\bsk \notin \calJ}}
\end{gather*}
Then $\APP'(\pm f,n) = \APP'(0,n)$, and
\begin{align*}
\MoveEqLeft{\max_{\pm} \norm[\calG]{\SOL(\pm f) - \APP'(\pm f,n)} =  \max_{\pm} \norm[\calG]{\SOL(\pm f) - \APP'(0,n)}} \\
& \ge \frac 12 \left [ \norm[\calG]{\SOL(f) - \APP'(0,n)} 
+ \norm[\calG]{- \SOL(f) - \APP'(0,n)}\right] \\
& \ge \norm[\calG]{\SOL(f)} 
= \Bignorm[\tau]{\bigl(\hf(\bsk)\bigr)_{\bsk \notin \calJ}}
    = \norm[\rho]{\left( \frac{\hf(\bsk)}{\lambda_{\bsk}} \right)_{\bsk \notin \calJ}} \,
    \norm[\rho']{\left( \lambda_{\bsk} \right)_{\bsk \notin \calJ}} \\
    & = R  \norm[\rho']{\left( \lambda_{\bsk} \right)_{\bsk \notin \calJ}}.
\end{align*}
The ordering of the $\lambda_{\bsk}$ implies that $\norm[\rho']{\left( \lambda_{\bsk} \right)_{\bsk \notin \calJ}}$ for arbitrary $\calJ$ can be no smaller than the case $\calJ = \{\bsk_1, \ldots, \bsk_n\}$.  This completes the proof.
\end{proof}

>>>>

\usepackage[utf8]{inputenc}%(only for the pdftex engine)
%\RequirePackage[no-math]{fontspec}[2017/03/31]%(only for the luatex or the xetex engine)
\usepackage[small]{dgruyter}
\usepackage{microtype}
\usepackage{amsmath,amssymb}
\usepackage{subcaption}
\usepackage{bbm}
\input{macros}

@@

\theoremstyle{dgthm}
\newtheorem{theorem}{Theorem}
\theoremstyle{dgthm}
\newtheorem{lemma}{Lemma}
\theoremstyle{dgthm}
\newtheorem{corollary}{Corollary}
\theoremstyle{dgthm}
\newtheorem{proposition}{Proposition}
\theoremstyle{dgdef}
\newtheorem{remark}{Remark}
\theoremstyle{definition}
\newtheorem{exmp}{Example}

\DeclareMathOperator{\SOL}{SOL}
\DeclareMathOperator{\APP}{APP}
\DeclareMathOperator{\ALG}{ALG}
\DeclareMathOperator{\ERR}{ERR}
\newcommand{\dataN}{\bigl(\hf(\bsk_i)\bigr)_{i=1}^n}
\newcommand{\dataNj}{\bigl(\hf(\bsk_i)\bigr)_{i=1}^{n_j}}
\newcommand{\dataNjd}{\bigl(\hf(\bsk_i)\bigr)_{i=1}^{n_{j^\dagger}}}
\newcommand{\ERRN}{\ERR\bigl(\dataN,n\bigr)}
\newcommand{\ERRNj}{\ERR\bigl(\dataNj,n_j\bigr)}
\newcommand{\ERRNjd}{\ERR\bigl(\dataNjd,n_{j^\dagger}\bigr)}
\DeclareMathOperator{\COST}{COST}
\DeclareMathOperator{\COMP}{COMP}
\newcommand{\hf}{\widehat{f}}
\newcommand{\hg}{\widehat{g}}
\newcommand{\tu}{\Tilde{u}}
\newcommand{\tv}{\Tilde{v}}
\newcommand{\tcalJ}{\widetilde{\calJ}}
\newcommand{\tcalK}{\widetilde{\calK}}
\newcommand{\tlambda}{\Tilde{\lambda}}
\newcommand{\tbbK}{\widetilde{\bbK}}
\newcommand{\lo}{\textup{lo}}
\newcommand{\up}{\textup{up}}
\newcommand{\E}{\textup{e}}
\def\abs#1{\ensuremath{\left \lvert #1 \right \rvert}}
\newcommand{\normabs}[1]{\ensuremath{\lvert #1 \rvert}}
\newcommand{\bigabs}[1]{\ensuremath{\bigl \lvert #1 \bigr \rvert}}
\newcommand{\Bigabs}[1]{\ensuremath{\Bigl \lvert #1 \Bigr \rvert}}
\newcommand{\biggabs}[1]{\ensuremath{\biggl \lvert #1 \biggr \rvert}}
\newcommand{\Biggabs}[1]{\ensuremath{\Biggl \lvert #1 \Biggr \rvert}}
\newcommand{\norm}[2][{}]{\ensuremath{\left \lVert #2 \right \rVert}_{#1}}
\newcommand{\normnorm}[2][{}]{\ensuremath{\lVert #2 \rVert}_{#1}}
\newcommand{\bignorm}[2][{}]{\ensuremath{\bigl \lVert #2 \bigr \rVert}_{#1}}
\newcommand{\Bignorm}[2][{}]{\ensuremath{\Bigl \lVert #2 \Bigr \rVert}_{#1}}
\newcommand{\biggnorm}[2][{}]{\ensuremath{\biggl \lVert #2 \biggr \rVert}_{#1}}
\newcommand{\ip}[3][{}]{\ensuremath{\left \langle #2, #3 \right \rangle_{#1}}}
\newcommand{\FredNote}[1]{{\color{blue}Fred: #1}}
\newcommand{\YuhanNote}[1]{{\color{magenta}Yuhan: #1}}
\newcommand{\PeterNote}[1]{{\color{orange}Peter: #1}}
\newcommand{\SimonNote}[1]{{\color{purple}Simon: #1}}
\newcommand{\tLambda}{\widetilde{\Lambda}}
%\newcommand{\HickernellFJ}{Hickernell}

@@

  \author[1]{Yuhan Ding}
  \author*[2]{Fred J. Hickernell}
  \author[3]{Peter Kritzer} 
  \author[4]{Simon Mak}
  \runningauthor{Y. Ding, F. J. Hickernell, P. Kritzer, and S. Mak}
  \affil[1]{Department of Mathematics, 
  Misericorida University, 301 Lake Street, Dallas, PA 18704 USA}
  \affil[2]{Department of Applied Mathematics, Illinois Institute of Technology, RE 220, 10 W.\ 32${}\text{nd}$ Street, Chicago, IL 60616 USA}
  \affil[3]{Johann Radon Institute for Computational and Applied Mathematics (RICAM), 
  Austrian Academy of Sciences, Altenbergerstr. 69, 4040 Linz, Austria}
  \affil[4]{H. Milton Stewart School of Industrial and Systems Engineering, Georgia Institute of Technology, 755 Ferst Drive, Atlanta, GA 30332 USA}
  \title{Adaptive Approximation for Multivariate Linear Problems with Inputs Lying in a Cone}
  \runningtitle{Adaptive Approximation}
  \subtitle{...}
  \abstract{...}
  %\keywords{...}
  %\classification[PACS]{...}
  %\communicated{...}
  %\dedication{...}
  %\received{...}
  %\accepted{...}
  %\journalname{...}
  %\journalyear{...}
  %\journalvolume{..}
  %\journalissue{..}
  \startpage{1}
  \aop
  %\DOI{...}

@@

In many situations, adaptive algorithms can be rigorously shown to perform \emph{essentially no better} than non-adaptive algorithms.  Yet, in practice adaptive algorithms are appreciated because they relieve the user from stipulating the computational effort required to achieve the desired accuracy.  The key to resolving this seeming contradiction is to construct a theory based on assumptions that favor adaptive algorithms. We do that here.

Adaptive algorithms infer the necessary computational effort based on the function data sampled.  Adaptive algorithms may perform better than non-adaptive algorithms if the set of input functions is not convex. We construct adaptive algorithms for general multivariate linear problems where the input functions lie in non-convex cones.  Our algorithms use series coefficients of the input function to construct an approximate solution that satisfies an absolute error tolerance.  We show our algorithms to be essentially optimal.  We derive conditions under which the problem is tractable.  In the remainder of this section we define the problem and essential notation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{An Illustrative Example}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Consider the case of approximating functions defined over $[-1,1]^d$, using a Chebyshev polynomial basis.  The input function is denoted $f$, and the solution is $\SOL(f) = f$ in this case
	\begin{align*}
	f &  = \sum_{\bsk \in \bbN_0^d} \widehat{f}(\bsk) u_\bsk =: \SOL(f) \\
    u_\bsk & := \prod_{\ell =1}^d \tu_{k_\ell} , & \tu_{k}(x) & := \begin{cases} 1, & k =0 ,\\
	x, & k=1,\\
	2x \tu_{k-1}(x)-\tu_{k-2}(x), & k = 2, 3, \ldots.
	\end{cases}
	\end{align*}
If one can observe the series coefficients, then $f$ can be approximated by a finite sum.  However, one should have some notion of which terms in the infinite series for $f$ are more important.  Let 
\begin{equation*}
    \lambda_\bsk := \prod_{\ell =1}^d \frac{1}{\max(1,k^r_\ell)}, \qquad \norm[\calF]{f} := \norm[2]{\left(\frac{\hf(\bsk)}{\lambda_{\bsk}}\right)_{\bsk \in \N_0^d}}.
\end{equation*}
By this definition, an input function with modest sized norm must have series coefficients that decay quickly enough as the degree of the polynomial increases.  The assumed decay is quicker for larger $r$, and this assumed decay governs the assumed smoothness of the input functions.

The weights, $\lambda_{\bsk}$, can be ordered,
\begin{equation} \label{DHKM:lambda_order}
    \lambda_{\bsk_1} \ge \lambda_{\bsk_2} \ge \cdots >0,
\end{equation}
which then implies an ordering of the wavenumbers.  Given this ordering, it is natural to approximate the solution using the first $n$ series coefficients as follows:
\begin{equation} \label{DHKM:Ex_APP_def}
   \APP(f,n) := \sum_{i=1}^n \hf(\bsk_i) u_{\bsk_i} \quad \forall n \in \N.
\end{equation}
The error of this approximation can be defined as 
\begin{gather*}
    \norm[\calG]{\SOL(f) - \APP(f,n)} = \norm[2]{\bigl(\hf(\bsk_i) \bigr)_{i=n+1}^{\infty}}, \\
    \text{where}\qquad
    \norm[\calG]{\sum_{\bsk \in \N_0^d} \hg(\bsk) u_{\bsk}} : = \norm[2]{\bigl(\hg(\bsk) \bigr)_{\bsk \in \N_0^d}}.
\end{gather*}

If one has a fixed data budget, $n$, then choosing approximation $\APP(f,n)$ is appropriate.  However, often one has an error tolerance, $\varepsilon$, for which one desires an algorithm, $\ALG(f,\varepsilon)$, which satisfies
\begin{equation*}
    \norm[\calG]{\SOL(f) - \ALG(f,\varepsilon)} \le \varepsilon.
\end{equation*}
Such an algorithm must have a rule of choosing the correct $n$---depending on $f$ and $\varepsilon$---so that $\ALG(f,\varepsilon) = \APP(f,n)$.  Choosing such a rule, choosing a set $\calC$ of input functions for which the rule is valid,  characterizing the information cost of $\ALG$, and determining whether $\ALG$ has optimal information cost, and understanding the dependence of this cost on the the number of input variables, $d$, as well as the error tolerance, $\varepsilon$, are all objectives of this chapter.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{General Linear Problem}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Our general linear problem is defined by a solution operator mapping an input function to an output, $\SOL:\calF \to \calG$.  As in the illustrative example above, the Banach spaces of inputs and outputs are defined by series expansions:
\begin{gather}
    \calF = \left \{f = \sum_{\bsk \in \mathbb{K}} \hf(\bsk) u_{\bsk} : \norm[\calF]{f} : = \norm[\rho]{\left( \frac{\hf(\bsk)}{\lambda_{\bsk}} \right)_{\bsk \in \mathbb{K}}} < \infty \right\}, \quad 1 \le \rho \le \infty, \\
    \calG = \left \{g = \sum_{\bsk \in \mathbb{K}} \hg(\bsk) v_{\bsk} : \norm[\calG]{g} : = \norm[\tau]{\bigl(  \hg(\bsk)  \bigr)_{\bsk \in \mathbb{K}}} < \infty \right\}, \quad 1 \le \tau \le \rho.
\end{gather}
Here, $\{u_{\bsk}\}_{\bsk \in \mathbb{K}}$ is a basis for the input Banach space $\calF$, and $\{v_{\bsk}\}_{\bsk \in \mathbb{K}}$ is a basis for the output Banach space $\calG$. These bases are defined to match the solution operator:
\begin{equation} \label{DHKM:basis_relate}
    \SOL(u_{\bsk}) = v_{\bsk} \qquad \forall \bsk \in \mathbb{K},
\end{equation}
where $\mathbb{K}$ is a general index set. The $\lambda_{\bsk}$ represent the importance of the series coefficients of the input function.  The larger $\lambda_{\bsk}$ is, the more important $\hf(\bsk)$ is.

Although this problem formulation is quite general in some aspects, condition \eqref{DHKM:basis_relate} is somewhat restrictive.  In theory, the choice of basis can be made via the singular value decomposition, but in practice, if the norms of $\calF$ and $\calG$ are typical, the bases satisfying \eqref{DHKM:basis_relate} may be difficult to identify explicitly.

To facilitate our derivations below, we establish the following lemma, which follows from H\"older's inequality:

\begin{lemma} \label{DHKM:Key_Lem}
Let $\calK$ be some proper or improper subset of $\bbK$. Moreover, let $\rho'$ be defined by the relation
\begin{equation*}
    \frac 1\rho + \frac 1 {\rho'} = \frac 1 \tau, \qquad \text{i.e., } \rho' = \frac{\rho \tau}{\rho - \tau},
\end{equation*}
so $\tau \le \rho' \le \infty$.  Let $\Lambda =  \bignorm[\rho']{\bigl(  \lambda_{\bsk}  \bigr)_{\bsk \in \calK}}$.  Then the following are true for $f = \sum_{\bsk \in \calK} \hf(\bsk) u_{\bsk}$:
\begin{equation}
\label{DHKM:SOL_ineq}
    \norm[\calG]{\SOL(f)} = \norm[\tau]{\bigl(\hf(\bsk) \bigr)_{\bsk \in \calK}} \le \norm[\calF]{f} \, \Lambda,
    \end{equation}
    \begin{multline}
    \label{DHKM:SOL_tight_ineq}
    \bigabs{\hf(\bsk)} = \begin{cases}
    \displaystyle 
    \frac{R \lambda_{\bsk}^{\rho'/\rho + 1}}{\Lambda^{\rho'/\rho}}, & \forall \bsk \in \calK, \quad  \rho'<\infty \\
    R \Lambda \delta_{\bsk,\bsk^*}, & \forall \bsk \in \calK, \ \bsk^* \in \calK \text{ satisfies } \lambda_{\bsk^*} = \Lambda, \quad \rho' = \infty,
    \end{cases}
   \\ 
    \implies  \ \ \norm[\calF]{f} = R \ \& \ \norm[\calG]{\SOL(f)} = R \Lambda.
    \end{multline}

\PeterNote{Please excuse me for asking stupid questions all the time, but I was wondering: if $\rho'=\infty$, how would the formula for the $\hf (\bsk)$ in (7) look like? We would have $\infty$ in the exponents.}
\FredNote{How does it look now?}
Equality \eqref{DHKM:SOL_tight_ineq} illustrates how inequality \eqref{DHKM:SOL_ineq} may be made tight.
\end{lemma}
\begin{proof}
We give the proof for $\rho' < \infty$.  The proof for $\rho' = \infty$ follows similiarly. 
The proof of inequality \eqref{DHKM:SOL_ineq} proceeds by applying H\"older's inequality:  
\begin{align}
    \label{DHKM:SOL_A}
    \norm[\calG]{\SOL(f)}  
    & = \biggnorm[\calG]{\sum_{\bsk \in \calK} \hf(\bsk) v_{\bsk}} = \norm[\tau]{\bigl(  \hf(\bsk)  \bigr)_{\bsk \in \calK}}  = \left [\sum_{\bsk \in \calK}  \left\lvert\frac{\hf(\bsk)}{\lambda_{\bsk}} \right\rvert^{\tau} \lambda_{\bsk}^{\tau} \right]^{1/\tau} \\
    \nonumber
    & \le \biggnorm[\rho]{\biggl(  \frac{\hf(\bsk)}{\lambda_{\bsk}}  \biggr)_{\bsk \in \calK}} \, \bignorm[\rho']{\bigl(  \lambda_{\bsk}  \bigr)_{\bsk \in \calK}} \qquad \text{since }\frac 1\rho + \frac 1 {\rho'} = \frac 1 \tau \\
    \nonumber
    & = \norm[\calF]{f} \, \Lambda.
\end{align}
Substituting the formula for $\bigabs{\hf(\bsk)}$ in \eqref{DHKM:SOL_tight_ineq} into equation \eqref{DHKM:SOL_A} and applying the relationship between $\rho$, $\rho'$, and $\tau$ yields
\begin{equation*}
       \norm[\calG]{\SOL(f)}  
    =  \frac{R \bignorm[\tau]{\bigl(  \lambda_{\bsk}^{\rho'/\rho + 1}  \bigr)_{\bsk \in \calK}}} {\Lambda^{\rho'/\rho}} 
    = \frac{R \bignorm[\rho']{\bigl(  \lambda_{\bsk}  \bigr)_{\bsk \in \calK}}^{\rho'/\rho + 1}}
    {\Lambda^{\rho'/\rho}} = R \Lambda.
\end{equation*}
Moreover,
\begin{equation*}
    \norm[\calF]{f}  
    = \norm[\rho]{\left( \frac{\hf(\bsk)}{\lambda_{\bsk}} \right)_{\bsk \in \calK}}
    = \frac{R \bignorm[\rho]{\bigl(  \lambda_{\bsk}^{\rho'/\rho}  \bigr)_{\bsk \in \calK}}}{\Lambda^{\rho'/\rho}} 
    = \frac{R \bignorm[\rho']{\bigl(  \lambda_{\bsk}  \bigr)_{\bsk \in \calK}}^{\rho'/\rho}}
    {\Lambda^{\rho'/\rho}} = R.
\end{equation*}
This completes the proof.
\end{proof} \

%\hspace{\parindent} 
Based on this lemma and taking $\calK = \bbK$, the norm of the solution operator can be expressed in terms of the $\lambda_\bsk$ as follows:
\begin{equation} \label{DHKM:SOLNorm}
    \norm[\calF \to \calG]{\SOL}  = \sup_{\norm[\calF]{f} \le 1} \norm[\calG]{\SOL(f)} = \bignorm[\rho']{\bigl(  \lambda_{\bsk}  \bigr)_{\bsk \in \mathbb{K}}}.
\end{equation}
The $\lambda_{\bsk}$ define the smoothness of the input space and are assumed to have a known order:
\begin{equation} \label{DHKM:lambda_order}
    \lambda_{\bsk_1} \ge \lambda_{\bsk_2} \ge \cdots >0.
\end{equation}
We assume that all $\lambda_{\bsk_i}$ are positive to avoid the trivial case where the problem can be solved by knowing a finite number of series coefficients.
An optimal approximation  based on $n$ series coefficients of the input function can be defined in terms of the series coefficients of the input function corresponding to the largest $\lambda_{\bsk}$ as follows:
\begin{equation} \label{DHKM:APP_def}
    \APP : \calF \times \N_0 \to \calG, \quad  \APP(f,0) = 0, \ \ \APP(f,n) := \sum_{i=1}^n \hf(\bsk_i) v_{\bsk_i} \ \forall n \in \N.
\end{equation}

@@

 Let $\APP'(0,n)$ depend on the series coefficients indexed by $\calJ  = \{\bsk'_1, \ldots, \bsk'_n\}$.  Use Lemma \ref{DHKM:Key_Lem} with $\calK = \bbK \setminus \calJ$ to choose $f$ to satisfy the following conditions:
\begin{gather*}
    \hf(\bsk'_1) = \cdots = \hf(\bsk'_n) = 0, \qquad \norm[\calF]{f} = R, \\ \Bignorm[\tau]{\bigl(\hf(\bsk)\bigr)_{\bsk \notin \calJ}}
    = \norm[\rho]{\left( \frac{\hf(\bsk)}{\lambda_{\bsk}} \right)_{\bsk \notin \calJ}} \,
    \norm[\rho']{\left( \lambda_{\bsk} \right)_{\bsk \notin \calJ}}.
\end{gather*}
Then $\APP'(\pm f,n) = \APP'(0,n)$, and
\begin{align*}
\MoveEqLeft{\max_{\pm} \norm[\calG]{\SOL(\pm f) - \APP'(\pm f,n)} =  \max_{\pm} \norm[\calG]{\SOL(\pm f) - \APP'(0,n)}} \\
& \ge \frac 12 \left [ \norm[\calG]{\SOL(f) - \APP'(0,n)} 
+ \norm[\calG]{- \SOL(f) - \APP'(0,n)}\right] \\
& \ge \norm[\calG]{\SOL(f)} 
 = R  \norm[\rho']{\left( \lambda_{\bsk} \right)_{\bsk \notin \calJ}} \qquad \text{by \eqref{DHKM:SOL_tight_ineq}}.
\end{align*}
The ordering of the $\lambda_{\bsk}$ implies that $\norm[\rho']{\left( \lambda_{\bsk} \right)_{\bsk \notin \calJ}}$ for arbitrary $\calJ$ can be no smaller than the case $\calJ = \{\bsk_1, \ldots, \bsk_n\}$.  This completes the proof.
\end{proof}

\hspace{\parindent} The decay rate of the approximation error in \eqref{DHKM:APP_errorBd} depends on the decay rate of the tail norm of the $\lambda_{\bsk_i}$.  This decay may be algebraic or exponential and will determine the complexity of the problem as a function of the error tolerance.
